Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-05 2024pipit/section/20-Motivation.tex
trigger index

\begin{tabbing}
  @MM@\= @MM@ \= @let index@ \= \kill
  @let@ trigger_fetch ($\textit{cfg}$: triggers) ($\textit{cycle}$: $\NN$) ($\textit{time}$: stream $\NN$): stream $\NN$ = \\
    \> $@rec @ \textit{index}.$ \\
    \> \> $@let @ \textit{inc} = \text{false} @ fby @ ((\text{time_mark}~\textit{cfg}~\textit{index} \le \textit{time}) \vee \neg (\text{enabled}~\textit{cfg}~\textit{index}~\textit{cycle})) @ in@$\\
    \> \> $@let @ \textit{index} = \mbox{count_when}_{\PSValid}~(\text{trigger_count}~\textit{cfg})~\textit{inc} @ in@$ \\
    \> \> $\xcheckP{\PSUnknown}{(\text{time_no_skips}~\textit{time} \implies \text{can_reach_next_active}~\textit{cfg}~\textit{cycle}~\textit{time}~\textit{index})}$; \\
    \> \> $\text{pose_forall}_{\PSValid}~(\forall \rawbind{i} \text{can_reach_next}~\textit{cfg}~\textit{cycle}~i)~\textit{index}$; \\
    \> \> $\textit{index}$\\
\end{tabbing}

The implementation of trigger_fetch defines a streaming function which takes a static trigger configuration, the cycle index (static for now), and the stream denoting the current time.
The static trigger configuration contains the constant triggers array, as well as proofs that the configuration is schedulable: for example, that the triggers are sorted by time-mark and have adequate gaps between their times.

@@

Finally, we use the pose_forall combinator to instantiate a non-streaming proof with a streaming argument: given a proof of $\forall \rawbind{(x: \tau)} p~x$ and a stream $s: \stream{\tau}$, we can infer that $p$ holds for all elements of the stream $s$.
We annotate the call with $\PSValid$ here to emphasise that the proof is not deferred, but we do not include this in practice.
The pose_forall function is in fact just a helper function implemented using the check, transition-system and bless functions already seen.

>>>>

\begin{tabbing}
  @MM@\= @MM@ \= @let index@ \= \kill
  @let@ trigger_index ($\textit{cfg}$: triggers) ($\textit{cycle}$: $\NN$) ($\textit{time}$: stream $\NN$): stream $\NN$ = \\
    \> $@rec @ \textit{index}.$ \\
    \> \> $@let @ \textit{inc} = \text{false} @ fby @ ((\text{time_mark}~\textit{cfg}~\textit{index} \le \textit{time}) \vee \neg (\text{enabled}~\textit{cfg}~\textit{index}~\textit{cycle})) @ in@$\\
    \> \> $@let @ \textit{index} = \mbox{count_when}_{\PSValid}~(\text{trigger_count}~\textit{cfg})~\textit{inc} @ in@$ \\
    \> \> $\xcheckP{\PSUnknown}{(\text{time_no_skips}~\textit{time} \implies \text{can_reach_next_active}~\textit{cfg}~\textit{cycle}~\textit{time}~\textit{index})}$; \\
    \> \> $\text{pose_forall}_{\PSValid}~(\forall \rawbind{i} \text{can_reach_next}~\textit{cfg}~\textit{cycle}~i)~\textit{index}$; \\
    \> \> $\textit{index}$\\
\end{tabbing}

The implementation of trigger_index defines a streaming function which takes a static trigger configuration, the cycle index (static for now), and the stream denoting the current time.
The static trigger configuration contains the constant triggers array, as well as proofs that the configuration describes a valid schedule: for example, that the triggers are sorted by time-mark and have adequate gaps between their times.

@@

We then use the pose_forall combinator to instantiate a non-streaming proof with a streaming argument: given a proof of $(\forall \rawbind{(x: \tau)} p~x)$ and a stream $(s: \stream{\tau})$, we can infer that $p$ holds for all elements of the stream $s$.
We annotate the call with $\PSValid$ here to emphasise that the proof is not deferred, but we do not include this in practice.
The pose_forall function is in fact just a helper function implemented using the check, transition-system and bless functions already seen.
The property we state here is that if the trigger at the current index is active, then there is a sufficient time gap to be able to reach the next active trigger after the current index.

After explicitly instantiating this lemma, the streaming invariant can be proved by induction on the transition system; the proof script is almost identical to that of count_when.
